---
title: "H7"
author: "ITS ANONOMUS OKAU"
date: "7/11/22"
output: html_document
---

```{r}
pacman::p_load(tidyverse, rvest)
```

> I decided to scrape some data from cyclingstage.com about the bike race Vuelta. I'm interested in finding out whether the stages has gotten longer over the years. Or maybe they got shorter at some point, because people stopped doping? Or maybe the first stages are longer than the last ones, because the riders have more energy in the beginning?

```{r scrape-function}
# This function is kindly provided by Adela
scrape <- function(website){
	url <- read_html(website)
	annual_table <- url %>% 
 			html_nodes("table") %>% # lucky for me, all my data is in a table
 			html_table()
  annual_table <- do.call(cbind,unlist(annual_table, recursive = FALSE))
 }
```

```{r loop}
# Create a container for the results
master_table=NULL 

# Loop over the years to iterate the scraping
for (year in 2014:2020){  
	url <- "https://www.cyclingstage.com/vuelta-" # first part of URL
	website <- paste0(url,year,"-route") # i have to glue the year in the middle of the URL
	annual_table <- scrape(website) # using the function
	master_table <- rbind(master_table, annual_table) # append data
}
master_table <- as_tibble(master_table) # make into tibble
```

```{r cleaning}
vuelta <- master_table %>% # here comes some cleaning
  rename(desc = `Start and finish`) %>% # renaming a weird column
  filter(desc != "Rest day" & desc != "rest day") %>% # getting rid of rest days
  separate(col = desc, into = c("year", "desc"), sep = 11) %>% # splitting columns to get the year
  separate(col = year, into = c(NA, "year"), sep = 7) # removing the "vuelta" part of the year

vuelta$Length <- str_remove(vuelta$Length, "km") # removing km from Length column
vuelta$Length <- str_replace(vuelta$Length, ",", ".") # replacing , with .
vuelta$Length <- as.numeric(vuelta$Length) # making Length numeric
vuelta$Nr <- as.numeric(vuelta$Nr) # making stage number numeric
```

```{r plotting}
vuelta %>% 
  ggplot(aes(x = Nr, y = Length, color = year)) + 
  geom_point() +
  geom_hline(yintercept = mean(vuelta$Length))
```


> Well, from here, nothing looks interesting. No obvious changes at least. I did plot the mean length of the stages, and we see that some stages are very short. Probably time trials? These are marked "TTT" or "ITT" in the dataframe. Let's have a look:

```{r}
vuelta$TT <- ifelse(vuelta$Type == "TTT" | vuelta$Type == "ITT", "time trial", "other") # making a new column marking timetrials

vuelta %>% 
  ggplot(aes(x = Nr, y = Length, color = TT)) + 
  geom_point() +
  geom_hline(yintercept = mean(vuelta$Length))
```

> Yup! The timetrials are always short!